Exploring the conformational dynamics of alanine dipeptide in solution subjected to an external electric field: A nonequilibrium molecular dynamics simulation

### 2013-11-19 writedown/npt-neq/alanine.tex
final version before submission to jctc

\title{Exploring the conformational dynamics of alanine dipeptide in solution subjected to an external electric field: A nonequilibrium molecular dynamics simulation}
% \title{The non-equilibrium molecular dynamics simulation of the alanine dipeptide under electric field}
\author{Han Wang}
\email{han.wang@fu-berlin.de}
\affiliation{Institute for Mathematics, Freie Universit\"at Berlin, Germany}
\author{Christof Sch\"utte}
\affiliation{Institute for Mathematics, Freie Universit\"at Berlin, Germany}
\affiliation{Zuse Institute Berlin, Germany}
\author{Giovanni Ciccotti}
% \affiliation{Department of Physics, University of Roma "La Sapienza", Italy}
\affiliation{School of Physics, Room 302B UCD-EMSC, University College Dublin, Belfield, Dublin 4, Ireland}
\affiliation{Dipartimento di Fisica, Università “La Sapienza” and CNISM, Piazzale Aldo Moro 5, 00185 Roma, Italy}
\author{Luigi Delle Site}
\affiliation{Institute for Mathematics, Freie Universit\"at Berlin, Germany}

\begin{abstract}
In this paper, we investigate the conformational dynamics of alanine
   dipeptide under an external electric field by 
   nonequilibrium molecular dynamics simulation.  We consider
   the case of a constant and of an oscillatory field. In this context we
propose a procedure to
   implement the temperature control, which removes the irrelevant thermal
effects of the field. For the constant field different time-scales are
   identified  in the conformational, dipole moment, and orientational
   dynamics. Moreover, we prove that the solvent structure only
   marginally changes when the external field is switched on.  In the
   case of oscillatory field, the conformational changes are shown to be as
   strong as in the previous case, and non-trivial nonequilibrium
   circular paths in the conformation space are revealed by calculating the
   integrated net probability fluxes.

\end{abstract}

@@

\section{Methodology}
\subsection{Nonequilibrium molecular dynamics simulation}
In this section we briefly review the dynamical approach to nonequilibrium
molecular dynamics (D-NEMD) \cite{ciccotti1975direct, ciccotti1979thought,
  orlandini2011hydrodynamics, orlandini2011hydrodynamics-01}.
In the following we denote the macroscopic observable by $O(t)$. If at $t$
the configurational probability distribution is $\rho(\vect x, t)$, where
$\vect x$ is the phase space variable, then the observable can be
expressed by
\begin{align}\label{eqn:tmp1}
  O(t) = \int d\vect x\, \hat O(\vect x)\rho(\vect x, t)  = \langle \hat O(\vect x), \rho(\vect x, t)\rangle,
\end{align}
we refer to $\hat O (\vect x)$ as the microscopic observable, 
which is measured at each point in the phase space $\vect x$.
Here we always assume that the initial probability distribution
$\rho(\vect x, 0)$ is known. In particular, in our case,
it is identical to the equilibrium distribution of the system
without EF.
The bracket on the right hand side of Eq.~\eqref{eqn:tmp1} denotes, as usual, the inner product in the
phase space.  We assume that the dynamics of the system is governed by the
Hamiltonian equation, i.e. $\dot {\vect x} = J \cdot \nabla_{\vect x}
\mathcal H(\vect x, t)$, where $\mathcal H$ is the time-dependent Hamiltonian,
and $J$ is
the symplectic matrix. The Liouville equation for the probability
distribution writes:
\begin{align}\label{eqn:tmp2}
  \frac{\partial \rho(\vect x, t)}{\partial t} = - iL(t) \rho(\vect x, t),
\end{align}
where $iL(t) = \{\cdot, \mathcal H\}$ is the Liouville operator.
Eq.~\eqref{eqn:tmp2}
can be formally solved by $\rho(\vect x, t) = U^\dagger(t,0) \rho(\vect x, 0)$,
\recheck{where $U^\dagger(t,0) = \mathcal T\exp\{-i \int_0^t d t' L(t')\}$, and $\mathcal T$ is the time ordering operator.}
On the other hand,
\begin{align}
  \frac{d \hat O(\vect x(t))}{dt} = \nabla_{\vect x}\hat O\cdot \dot{\vect x}
  = \nabla_{\vect x}\hat O\cdot J\cdot \nabla_{\vect x}\mathcal H
  = iL(t) \hat O (\vect x(t)).
\end{align}
This equation can be formally solved by
$\hat O(\vect x(t)) = U(t,0)\,\hat O(\vect x(0))$, therefore,
\begin{align}\nonumber
  O(t) & = \langle \hat O(\vect x), \rho(\vect x, t)\rangle
  = \langle \hat O(\vect x), U^\dagger(t,0)\, \rho(\vect x, 0)\rangle
  = \langle U(t,0)\,\hat O(\vect x), \rho(\vect x, 0)\rangle\\\label{eqn:tmp4}
  &= \langle \hat O(\vect x(t)), \rho(\vect x, 0)\rangle.
\end{align}
Since we assume that the system starts from the equilibrium distribution (without EF)
 Eq.~\eqref{eqn:tmp4} expresses the fact that the 
observable $O(t)$ calculated in a situation of nonequilibrium is equal to the ensemble average of
the microscopic observable computed along trajectories, starting from
initial configurations sampled from an \emph{equilibrium}
distribution. In practical terms we proceed by first running an equilibrium MD simulation
to generate a sample of configurations. Next we employ these configurations as
initial configurations and, for each, the full Hamiltonian dynamics is integrated until
time $t$.
In this paper, these trajectories are called \emph{branching
  trajectories}.
% {\bf EITHER WE PUT THE PICTURE OR REMOVE THIS SENTENCE {(see Fig.\ref{branching} for a pictorial representation)}}. 
Finally, the value of the macroscopic observable at time $t$ is
calculated by averaging the microscopic observable measured at each time $t$ along each branching trajectory.
Stochastic dynamics (e.g.~Langevin dynamics) can be handled analogously.

@@

\begin{figure}
  \centering
  \includegraphics[width=1.00\textwidth]{fig-all.eps}
  % \raisebox{0.2\height}{\includegraphics[width=0.22\textwidth]{c-2.eps}}
  % \includegraphics[width=0.37\textwidth]{fig-begin-1.eps}
  % \includegraphics[width=0.37\textwidth]{fig-end-1.eps}
  \caption{
    The probability density of the conformations of the alanine dipeptide plotted on the Ramachandran plot.
    (a): Two dihedral angles
    $\phi$ and $\psi$ are used to represent the molecular conformations of the alanine dipeptide.
    (b): The equilibrium probability density 
    at $t=0$~\textsf{ps}. (c): The probability density of equilibrium at $t=1000$~\textsf{ps}, resulting from the application of a constant EF of 1~V/nm. Darker color
    indicates higher probability.
    Plots are divided into 5 subregions corresponding
    to 5 molecular conformations:
    $A_1 = \{(\phi, \psi) | -134^\circ \leq \phi <  13^\circ, -125^\circ \leq \psi < 74^\circ\}$,
    $A_2 = \{(\phi, \psi) |  128^\circ \leq \phi < 226^\circ, -125^\circ \leq \psi < 74^\circ\}$,
    $B_1 = \{(\phi, \psi) | -110^\circ \leq \phi <  13^\circ,   74^\circ \leq \psi < 235^\circ\}$,
    $B_2 = \{(\phi, \psi) |  128^\circ \leq \phi < 250^\circ,   74^\circ \leq \psi < 235^\circ\}$,
    $C   = \{(\phi, \psi) |   13^\circ \leq \phi < 128^\circ, -180^\circ \leq \psi < 180^\circ\}$.
    \recheck{Considering the angular periodicity, The angle, for example, $226^\circ$ corresponds to $-134^\circ$
    on the plot (b) and (c).}
    % Considering the angular periodicity, if we write, for example,
    % $128^\circ \leq \phi <-134^\circ$, it implies that $\phi$ is in a range of $98$ degrees, between $128^\circ$ and $226^\circ$ ($226^\circ$ correspond to $-134^\circ$).
    This division is the same for both plot (b) and (c).
  }
  \label{fig:tmp4}
\end{figure}

@@

From Fig.~\ref{fig:tmp4}, it is evident that the probability with which a
conformation occurs changes when the external EF is applied.
The most evident case is conformation $C$; in fact it is {almost} not present before the external field is applied but appears in a clear way as a response to the action of the EF. Following this line of thought, we are interested to study the change of conformation under the action of EF. Specifically, for each conformational change, we will analyze the relation between the structural relaxation of the molecule and its corresponding time scale. 
The starting point is the calculation of the probability
of each conformation:
\recheck{
\begin{align}
  P_I(t) = \mathbb P ( (\phi_t,\psi_t) \in I), \quad  I \in \{A_1, A_2, B_1, B_2, C\},
\end{align}
}
Notice that the probability is time dependent, being an observable in nonequilibrium situation.
Next,
we consider the net probability flux from conformation $I$ to $J$, defined by:
\recheck{
\begin{align}\nonumber
  F_{J,I}(t) & = \frac1{\Delta t}\,
  \bigg[\,
  \mathbb P \big( (\phi_{t-\Delta t},\psi_{t-\Delta t}) \in J,\, (\phi_t, \psi_t) \in I \,\big) 
  -
  \mathbb P \big( (\phi_{t-\Delta t}, \psi_{t-\Delta t}) \in I,\, (\phi_t, \psi_t) \in J \,\big)
  \bigg], \\\label{eqn:tmp7}
  & J,I \in \{A_1, A_2, B_1, B_2, C\}.
\end{align}}
\recheck{\textit{I cannot read what Giovanni wrote on the top of page 9.}}
Where $\mathbb P\big( (\phi_{t-\Delta t}, \psi_{t-\Delta t}) \in J,\, (\phi_t, \psi_t) \in I\,\big)$
is the joint probability of the alanine being in conformation $J$ at
time $(t-\Delta t)$ \emph{and} being in conformation $I$ at time $t$.
The same definition applies to
$\mathbb P \big( (\phi_{t-\Delta t}, \psi_{t-\Delta t}) \in I,\, (\phi_t, \psi_t) \in J\,\big)$.
Here the time interval $\Delta t$ should be small enough compared to the
time scale of conformational dynamics,
so that the changes in the conformation probabilities
can be treated in linear approximation.
At the same time,  $\Delta t$ should be also large enough compared with
the timestep of the MD integrator, 
so that the quantities of Eq.~\eqref{eqn:tmp7} are well-defined, and can be
estimated with sufficient numerical accuracy (see also Ref.~\cite{schuette2011markov}).
A positive value of $F_{J,I}(t)$ indicates that the
net flux goes from $J$ to $I$,
while a negative value indicates a net flux from $I$ to $J$.
Through $F_{J,I}(t)$, the analysis of the nonequilibrium process 
is projected onto the analysis of the probabilistic link between discretized conformations $\{A_1, A_2, B_1, B_2, C\}$.
In the case of oscillatory EF, the behaviour of the net probability flux is also highly oscillatory, thus, in order to clearly identify the essential features of the molecular conformational changes, it is convenient to study the integrated net probability flux, defined as:
\begin{align}\label{eqn:tmp7a}
  Q_{J,I} (t) = \int_0^t F_{J,I}(\tau)d \tau,
\end{align}
which expresses the cumulative effects due to the action of EF on the conformations of the molecule.

@@

\subsection{Molecular dipole response}
The fact that our external perturbation corresponds to the action of an electric field, naturally leads to the question of how the behaviour of the electric dipole moment of the molecule is influenced by the direct interaction with EF, and, in turn, how this is related to the overall conformational behaviour of the molecule reported in the previous section.
\begin{figure}
  \centering
  \includegraphics[width=0.45\textwidth]{fig-dipol-order-conf.eps}
  \caption{The
    $x$-component of the molecular dipole moment, and the orientational order parameter
    of the alanine molecule as a
    function of time. \recheck{The solid line represents the dipole moment,
    while the dashed line represents the orientational order parameter.}
    The left vertical axis is for the dipole moment, while
    the right is for the orientational order parameter.
  }
  \label{fig:tmp3}
\end{figure}

The dipole moment of the alanine dipeptide is defined by
\begin{align}\label{eqn:tmp8}
  \boldsymbol\mu_{\textrm{alanine}}(t) =
  \bigg\langle
  \sum_{i\in \{\textrm{atoms of alanine}\}}
  q_i\, \vect r_i(t)
  \bigg\rangle,
\end{align}
\recheck{where $q_i$ denotes the partial charge of any point particle $i$ defining our model of alnine,
the molecule being neutral, i.e.~$\sum_iq_i = 0$.}
$\vect r_i$ denotes the position of the particle $i$.
Fig.~\ref{fig:tmp3}
shows that the $x$-component of the dipole moment 
reaches 85\% of its maximum value in only 20~ps, which is comparable to
the warm-up time $t_{warm}$, and is 25 times \recheck{\textit{25 time should be correct, isn't it?}}
smaller than the slowest time scale of the conformational relaxation.
Then in the following 400~ps, the dipole slowly relaxes to its 
maximum value, that is, 
6.8~Debye, and  corresponds to an energy
($\boldsymbol\mu_{\textrm{alanine}} \cdot \vect E$)
of c.a.~$-13.6$~kJ/mol.
We calculate the averaged dipole moment of different conformations
(defined by taking average in Eqn.~\eqref{eqn:tmp8}
only for a certain conformation)
under the constant EF, and they are 6.8, 6.0, 5.1, 3.1 and 7.1~Debye
for $A_1$, $A_2$, $B_1$, $B_2$ and $C$, respectively.
The dipole energy difference between 3.1 and 7.1~Debye is roughly 8.0~kJ/mol.
Under a constant EF, the system will be likely to be driven towards those
conformations with higher dipole moment, because the
energy of the system will be lowered.  We suggest that this energy difference may be the reason why
we observe that the $\beta$-sheet conformations with lower dipole moment
are driven towards $\alpha$-helix conformations
$A_1$ and $C$, whose dipole moments are the highest
among all conformations. Moreover, the $B_2$ conformation vanishes because
its dipole is significantly lower than other conformations.
Comparing Fig.~\ref{fig:tmp3} with~\ref{fig:tmp5},
  the sharp increment of dipole moment before 20~ps is due to the quick vanishing of conformation
  $B_2$, while the slow increment of dipole moment until 400~ps is due to
  the slow migration from conformation $B_1$ to $A_1$ and $C$.
  % \textit{NOTE: here we want to discuss the relation between
  %   the conformation and the dipole moment. I rewote this
  %   sentence.}
% From 0 to 20~ps, the probability of
% lowest dipole conformation $B_2$ quickly
% vanishes. We observe a sharp increment of the average molecular dipole and finally, approaching  500~ps, the conformation slowly migrates from
% $B_1$ to $A_1$ and $C$. Moreover, since the dipole moment $A_1$ and $C$
% is higher than $B_1$, the average molecular dipole moment slowly increases
% to the nonequilibrium steady value.

@@

Fig.~\ref{fig:tmp9} shows the probability of the
conformations against time for periodically oscillatory EF,
while  
Fig.~\ref{fig:tmp10} presents the integrated net probability fluxes
(defined by Eqn.~\eqref{eqn:tmp7a}) between
conformations. Here the probability fluxes are not reported
because the profiles are highly oscillating and they would not offer 
a better information than the one that can be obtained from the integrated probability flux.
Coming back to Fig.~\ref{fig:tmp9}, for all periods investigated,
the observed time-dependent probabilities  in $A_1$, $A_2$, $B_1$ and $B_2$ are
highly oscillating and the average value over time cycles does not change
considerably w.r.t.~time.
However, the probability in conformation $C$ significantly
increases to approximately 0.17 for $\period =10$~ps, 0.27 for  $\period =40$~ps, and
0.25 for $\period =200$~ps.
In the case of $\period =10$ and $40$~ps, {the probability of conformation $C$}
reaches the steady value in around 300~ps, while it requires about 1200~ps
for the $\period =200$~ps case. 
% \bluec{\textit{Here I do not agree with you to remove the above blue discussion,
%   because the time scales (also the oscillatory EF case) are
%   the main results of the current paper, which should be presented. It is only one sentence, no one would be panic.}}
The notable increment of probability of $A_1$ and
  the vanishing of $B_1$ in the constant EF case (see Fig.~\ref{fig:tmp5})
are not observed in the oscillatory EF case.
\begin{figure}
  \centering
  \includegraphics[width=0.7\textwidth]{graph-t0010-2.eps}
  \caption{Schematic plot of the main probability flux between conformations
    for periodically oscillatory EF, $\period =10$~ps. The thickness 
    of the arrows approximately presents the strength of the flux.
    The numbers near the arrows indicates the strength of the averaged
    probability flux, the unit of which is $10^{-3}\times\textrm{ps}^{-1}$.
    % {\bf LUIGI COMMENT: THE PICTURE MUST BE MUCH MUCH LAREGER SO THAT ONE SEE THE DIFFERENCE IN THE STRUCTURES. mOREOVER, THEY HAVE AN INTEREST ONLY THEY CAN BE RELATED TO MORE GENERAL STRUCTURES. wHAT IS NEEDED IS THAT CLOSE TO EACH MOLECULE ONE SHOULD WRITE ``$\alpha$-helix-like'' ``$\beta$-sheet-like. THE REST, gIOVANNI IS RIGHT, DOES NOT NEED TO BE EXPALINED IN DETAIL BECAUSE IT IS NOT REALLY INTERESTING AND i CUT A LOT.}
  }
  \label{fig:tmp11}
\end{figure}
In comparison to the static EF case, where the probability flux
vanishes when the system reaches steady state (Fig.~\ref{fig:tmp6}), 
Fig.~\ref{fig:tmp10} shows increasing
integrated probability net fluxes, which imply some lasting and directional
fluxes among the conformations (see Fig.~\ref{fig:tmp9}).
Fig.~\ref{fig:tmp11} presents
a schematic plot of the main probability fluxes for $\period =10$~ps.
The thickness of the arrows and the numbers nearby indicate
the steady value of average net fluxes over periods.
% the average amount of the net fluxes.
%\recheck{This shows an interesting phenomenon: although
%the oscillatory EF vanishes by time-averaging over a period,
%it generates directional probability fluxes among the conformations.}

>>>>

\title{Exploring the conformational dynamics of alanine dipeptide in solution subjected to an external electric field: A nonequilibrium molecular dynamics simulation}
\author{Han Wang}
\email{han.wang@fu-berlin.de}
\affiliation{Institute for Mathematics, Freie Universit\"at Berlin, Germany}
\author{Christof Sch\"utte}
\affiliation{Institute for Mathematics, Freie Universit\"at Berlin, Germany}
\affiliation{Zuse Institute Berlin, Germany}
\author{Giovanni Ciccotti}
% \affiliation{Department of Physics, University of Roma "La Sapienza", Italy}
\affiliation{School of Physics, Room 302B UCD-EMSC, University College Dublin, Belfield, Dublin 4, Ireland}
\affiliation{Dipartimento di Fisica, Universit\`a ``La Sapienza'' and CNISM, Piazzale Aldo Moro 5, 00185 Roma, Italy}
\author{Luigi Delle Site}
\affiliation{Institute for Mathematics, Freie Universit\"at Berlin, Germany}

\begin{abstract}
In this paper, we investigate the conformational dynamics of alanine
   dipeptide under an external electric field by 
   nonequilibrium molecular dynamics simulation.  We consider
   the case of a constant and of an oscillatory field. In this context we
propose a procedure to
   implement the temperature control, which removes the irrelevant thermal
effects of the field. For the constant field different time-scales are
   identified  in the conformational, dipole moment, and orientational
   dynamics. Moreover, we prove that the solvent structure only
   marginally changes when the external field is switched on.  In the
   case of oscillatory field, the conformational changes are shown to be as
   strong as in the previous case, and non-trivial nonequilibrium
   circular paths in the conformation space are revealed by calculating the
   integrated net probability fluxes.
  % In this paper, we investigate the conformational dynamics of alanine dipeptide under an external electric field by nonequilibrium molecular dynamics simulation.  We consider the case of a constant and of an oscillatory field. In this context we propose a procedure to implement the temperature control, which removes the irrelevant thermal effects of the field. For the constant field different time-scales are identified  in the conformational, dipole moment, and orientational dynamics. Moreover, we prove that the solvent structure only marginally changes when the external field is switched on.  In the case of oscillatory field, the conformational changes are shown to be as strong as in the previous case, and non-trivial nonequilibrium circular paths in the conformation space are revealed by calculating the integrated net probability fluxes.
\end{abstract}

@@

\section{Methodology}
\subsection{Nonequilibrium molecular dynamics simulation}
In this section we briefly review the dynamical approach to nonequilibrium
molecular dynamics (D-NEMD) \cite{ciccotti1975direct, ciccotti1979thought,
  orlandini2011hydrodynamics, orlandini2011hydrodynamics-01}.
In the following we denote the macroscopic observable by $O(t)$. If at $t$
the configurational probability distribution is $\rho(\vect x, t)$, where
$\vect x$ is the phase space variable, then the observable can be
expressed by
\begin{align}\label{eqn:tmp1}
  O(t) = \int d\vect x\, \hat O(\vect x)\rho(\vect x, t)  = \langle \hat O(\vect x), \rho(\vect x, t)\rangle,
\end{align}
we refer to $\hat O (\vect x)$ as the microscopic observable, 
which is measured at each point in the phase space $\vect x$.
Here we always assume that the initial probability distribution
$\rho(\vect x, 0)$ is known. In particular, in our case,
it is identical to the equilibrium distribution of the system
without EF.
The bracket on the right hand side of Eq.~\eqref{eqn:tmp1} denotes, as usual, the inner product in the
phase space.  We assume that the dynamics of the system is governed by the
Hamiltonian equation, i.e. $\dot {\vect x} = J \cdot \nabla_{\vect x}
\mathcal H(\vect x, t)$, where $\mathcal H$ is the time-dependent Hamiltonian,
and $J$ is
the symplectic matrix. The Liouville equation for the probability
distribution writes:
\begin{align}\label{eqn:tmp2}
  \frac{\partial \rho(\vect x, t)}{\partial t} = - iL(t) \rho(\vect x, t),
\end{align}
where $iL(t) = \{\cdot, \mathcal H\}$ is the Liouville operator.
Eq.~\eqref{eqn:tmp2}
can be formally solved by $\rho(\vect x, t) = U^\dagger(t,0) \rho(\vect x, 0)$,
{where $U^\dagger(t,0) = \mathcal T\exp\{-i \int_0^t d t' L(t')\}$, and $\mathcal T$ is the time ordering operator.}
On the other hand,
\begin{align}
  \frac{d \hat O(\vect x(t))}{dt} = \nabla_{\vect x}\hat O\cdot \dot{\vect x}
  = \nabla_{\vect x}\hat O\cdot J\cdot \nabla_{\vect x}\mathcal H
  = iL(t) \hat O (\vect x(t)).
\end{align}
This equation can be formally solved by
$\hat O(\vect x(t)) = U(t,0)\,\hat O(\vect x(0))$, therefore,
\begin{align}\nonumber
  O(t) & = \langle \hat O(\vect x), \rho(\vect x, t)\rangle
  = \langle \hat O(\vect x), U^\dagger(t,0)\, \rho(\vect x, 0)\rangle
  = \langle U(t,0)\,\hat O(\vect x), \rho(\vect x, 0)\rangle\\\label{eqn:tmp4}
  &= \langle \hat O(\vect x(t)), \rho(\vect x, 0)\rangle.
\end{align}
Since we assume that the system starts from the equilibrium distribution (without EF)
 Eq.~\eqref{eqn:tmp4} expresses the fact that the 
observable $O(t)$ calculated in a situation of nonequilibrium is equal to the ensemble average of
the microscopic observable computed along trajectories, starting from
initial configurations sampled from an \emph{equilibrium}
distribution. In practical terms we proceed by first running an equilibrium MD simulation
to generate a sample of configurations. Next we employ these configurations as
initial configurations and, for each, the full Hamiltonian dynamics is integrated until
time $t$.
In this paper, these trajectories are called \emph{branching
  trajectories}.
% {\bf EITHER WE PUT THE PICTURE OR REMOVE THIS SENTENCE {(see Fig.\ref{branching} for a pictorial representation)}}. 
Finally, the value of the macroscopic observable at time $t$ is
calculated by averaging the microscopic observable measured at each time $t$ along each branching trajectory.
Stochastic dynamics (e.g.~Langevin dynamics) can be handled analogously.

@@

\begin{figure}
  \centering
  \includegraphics[width=1.00\textwidth]{fig-all.eps}
  % \raisebox{0.2\height}{\includegraphics[width=0.22\textwidth]{c-2.eps}}
  % \includegraphics[width=0.37\textwidth]{fig-begin-1.eps}
  % \includegraphics[width=0.37\textwidth]{fig-end-1.eps}
  \caption{
    The probability density of the conformations of the alanine dipeptide plotted on the Ramachandran plot.
    (a): Two dihedral angles
    $\phi$ and $\psi$ are used to represent the molecular conformations of the alanine dipeptide.
    (b): The equilibrium probability density 
    at $t=0$~\textsf{ps}. (c): The probability density of equilibrium at $t=1000$~\textsf{ps}, resulting from the application of a constant EF of 1~V/nm. Darker color
    indicates higher probability.
    Plots are divided into 5 subregions corresponding
    to 5 molecular conformations:
    $A_1 = \{(\phi, \psi) | -134^\circ \leq \phi <  13^\circ, -125^\circ \leq \psi < 74^\circ\}$,
    $A_2 = \{(\phi, \psi) |  128^\circ \leq \phi < 226^\circ, -125^\circ \leq \psi < 74^\circ\}$,
    $B_1 = \{(\phi, \psi) | -110^\circ \leq \phi <  13^\circ,   74^\circ \leq \psi < 235^\circ\}$,
    $B_2 = \{(\phi, \psi) |  128^\circ \leq \phi < 250^\circ,   74^\circ \leq \psi < 235^\circ\}$,
    $C   = \{(\phi, \psi) |   13^\circ \leq \phi < 128^\circ, -180^\circ \leq \psi < 180^\circ\}$.
    Considering the angular periodicity, The angle, for example, $226^\circ$ corresponds to $-134^\circ$
    on the plot (b) and (c).
    % Considering the angular periodicity, if we write, for example,
    % $128^\circ \leq \phi <-134^\circ$, it implies that $\phi$ is in a range of $98$ degrees, between $128^\circ$ and $226^\circ$ ($226^\circ$ correspond to $-134^\circ$).
    This division is the same for both plot (b) and (c).
  }
  \label{fig:tmp4}
\end{figure}

@@

From Fig.~\ref{fig:tmp4}, it is evident that the probability with which a
conformation occurs changes when the external EF is applied.
The most evident case is conformation $C$; in fact it is {almost} not present before the external field is applied but appears in a clear way as a response to the action of the EF. Following this line of thought, we are interested to study the change of conformation under the action of EF. Specifically, for each conformational change, we will analyze the relation between the structural relaxation of the molecule and its corresponding time scale. 
The starting point is the calculation of the probability
of each conformation:
\begin{align}
  P_I(t) = \mathbb P ( (\phi_t,\psi_t) \in I), \quad  I \in \{A_1, A_2, B_1, B_2, C\},
\end{align}
Notice that the probability is time dependent, being an observable in nonequilibrium situation.
Next,
we consider the net probability flux from conformation $I$ to $J$, defined by:
\begin{align}\nonumber
  F_{J,I}(t) & = \frac1{\Delta t}\,
  \bigg[\,
  \mathbb P \big( (\phi_{t-\Delta t},\psi_{t-\Delta t}) \in J,\, (\phi_t, \psi_t) \in I \,\big) 
  -
  \mathbb P \big( (\phi_{t-\Delta t}, \psi_{t-\Delta t}) \in I,\, (\phi_t, \psi_t) \in J \,\big)
  \bigg], \\\label{eqn:tmp7}
  & J,I \in \{A_1, A_2, B_1, B_2, C\}.
\end{align}
% {\bf Here for $\lim_{\Delta \to 0}$ is intended that in the simulation the time interval is taken large enough so that we have  enough events and thus stable numbers (see also Ref.~\cite{schuette2011markov})}.
Where $\mathbb P\big( (\phi_{t-\Delta t}, \psi_{t-\Delta t}) \in J,\, (\phi_t, \psi_t) \in I\,\big)$
is the joint probability of the alanine being in conformation $J$ at
time $(t-\Delta t)$ \emph{and} being in conformation $I$ at time $t$.
The same definition applies to
$\mathbb P \big( (\phi_{t-\Delta t}, \psi_{t-\Delta t}) \in I,\, (\phi_t, \psi_t) \in J\,\big)$.
Here the time interval $\Delta t$ should be small enough compared to the
time scale of conformational dynamics,
so that the changes in the conformation probabilities
can be treated in linear approximation.
At the same time,  $\Delta t$ should be also large enough compared with
the timestep of the MD integrator, 
so that the quantities of Eq.~\eqref{eqn:tmp7} are well-defined, and can be
estimated with sufficient numerical accuracy (see also Ref.~\cite{schuette2011markov}).
A positive value of $F_{J,I}(t)$ indicates that the
net flux goes from $J$ to $I$,
while a negative value indicates a net flux from $I$ to $J$.
Through $F_{J,I}(t)$, the analysis of the nonequilibrium process 
is projected onto the analysis of the probabilistic link between discretized conformations $\{A_1, A_2, B_1, B_2, C\}$.
In the case of oscillatory EF, the behaviour of the net probability flux is also highly oscillatory, thus, in order to clearly identify the essential features of the molecular conformational changes, it is convenient to study the integrated net probability flux, defined as:
\begin{align}\label{eqn:tmp7a}
  Q_{J,I} (t) = \int_0^t F_{J,I}(\tau)d \tau,
\end{align}
which expresses the cumulative effects due to the action of EF on the conformations of the molecule.

@@

\subsection{Molecular dipole response}
The fact that our external perturbation corresponds to the action of an electric field, naturally leads to the question of how the behaviour of the electric dipole moment of the molecule is influenced by the direct interaction with EF, and, in turn, how this is related to the overall conformational behaviour of the molecule reported in the previous section.
\begin{figure}
  \centering
  \includegraphics[width=0.45\textwidth]{fig-dipol-order-conf.eps}
  \caption{The
    $x$-component of the molecular dipole moment, and the orientational order parameter
    of the alanine molecule as a
    function of time. The solid line represents the dipole moment,
    while the dashed line represents the orientational order parameter.
    The left vertical axis is for the dipole moment, while
    the right is for the orientational order parameter.
  }
  \label{fig:tmp3}
\end{figure}

The dipole moment of the alanine dipeptide is defined by
\begin{align}\label{eqn:tmp8}
  \boldsymbol\mu_{\textrm{alanine}}(t) =
  \bigg\langle
  \sum_{i\in \{\textrm{atoms of alanine}\}}
  q_i\, \vect r_i(t)
  \bigg\rangle,
\end{align}
where $q_i$ denotes the partial charge of any point particle $i$ defining our model of alanine,
the molecule being neutral, i.e.~$\sum_iq_i = 0$.
$\vect r_i$ denotes the position of the particle $i$.
Fig.~\ref{fig:tmp3}
shows that the $x$-component of the dipole moment 
reaches 85\% of its maximum value in only 20~ps, which is comparable to
the warm-up time $t_{warm}$, and is 25 times
smaller than the slowest time scale of the conformational relaxation.
Then in the following 400~ps, the dipole slowly relaxes to its 
maximum value, that is, 
6.8~Debye, and  corresponds to an energy
($\boldsymbol\mu_{\textrm{alanine}} \cdot \vect E$)
of c.a.~$-13.6$~kJ/mol.
We calculate the averaged dipole moment of different conformations
(defined by taking average in Eqn.~\eqref{eqn:tmp8}
only for a certain conformation)
under the constant EF, and they are 6.8, 6.0, 5.1, 3.1 and 7.1~Debye
for $A_1$, $A_2$, $B_1$, $B_2$ and $C$, respectively.
The dipole energy difference between 3.1 and 7.1~Debye is roughly 8.0~kJ/mol.
Under a constant EF, the system will be likely to be driven towards those
conformations with higher dipole moment, because the
energy of the system will be lowered.  We suggest that this energy difference may be the reason why
we observe that the $\beta$-sheet conformations with lower dipole moment
are driven towards $\alpha$-helix conformations
$A_1$ and $C$, whose dipole moments are the highest
among all conformations. Moreover, the $B_2$ conformation vanishes because
its dipole is significantly lower than other conformations.
Comparing Fig.~\ref{fig:tmp3} with~\ref{fig:tmp5},
  the sharp increment of dipole moment before 20~ps is due to the quick vanishing of conformation
  $B_2$, while the slow increment of dipole moment until 400~ps is due to
  the slow migration from conformation $B_1$ to $A_1$ and $C$.
  % \textit{NOTE: here we want to discuss the relation between
  %   the conformation and the dipole moment. I rewote this
  %   sentence.}
% From 0 to 20~ps, the probability of
% lowest dipole conformation $B_2$ quickly
% vanishes. We observe a sharp increment of the average molecular dipole and finally, approaching  500~ps, the conformation slowly migrates from
% $B_1$ to $A_1$ and $C$. Moreover, since the dipole moment $A_1$ and $C$
% is higher than $B_1$, the average molecular dipole moment slowly increases
% to the nonequilibrium steady value.

@@

Fig.~\ref{fig:tmp9} shows the probability of the
conformations against time for periodically oscillatory EF,
while  
Fig.~\ref{fig:tmp10} presents the integrated net probability fluxes
(defined by Eqn.~\eqref{eqn:tmp7a}) between
conformations. Here the probability fluxes are not reported
because the profiles are highly oscillating and they would not offer 
a better information than the one that can be obtained from the integrated probability flux.
Coming back to Fig.~\ref{fig:tmp9}, for all periods investigated,
the observed time-dependent probabilities  in $A_1$, $A_2$, $B_1$ and $B_2$ are
highly oscillating and the average value over time cycles does not change
considerably w.r.t.~time.
However, the probability in conformation $C$ significantly
increases to approximately 0.17 for $\period =10$~ps, 0.27 for  $\period =40$~ps, and
0.25 for $\period =200$~ps.
In the case of $\period =10$ and $40$~ps, {the probability of conformation $C$}
reaches the steady value in around 300~ps, while it requires about 1200~ps
for the $\period =200$~ps case. 
The notable increment of probability of $A_1$ and
  the vanishing of $B_1$ in the constant EF case (see Fig.~\ref{fig:tmp5})
are not observed in the oscillatory EF case.
\begin{figure}
  \centering
  \includegraphics[width=0.6\textwidth]{graph-t0010-2.eps}
  \caption{Schematic plot of the main probability flux between conformations
    for periodically oscillatory EF, $\period =10$~ps. The thickness 
    of the arrows approximately presents the strength of the flux.
    The numbers near the arrows indicates the strength of the averaged
    probability flux, the unit of which is $10^{-3}\times\textrm{ps}^{-1}$.
  }
  \label{fig:tmp11}
\end{figure}
In comparison to the static EF case, where the probability flux
vanishes when the system reaches steady state (Fig.~\ref{fig:tmp6}), 
Fig.~\ref{fig:tmp10} shows increasing
integrated probability net fluxes, which imply some lasting and directional
fluxes among the conformations (see Fig.~\ref{fig:tmp9}).
Fig.~\ref{fig:tmp11} presents
a schematic plot of the main probability fluxes for $\period =10$~ps.
The thickness of the arrows and the numbers nearby indicate
the steady value of average net fluxes over periods.
{
The relevance of Fig.~\ref{fig:tmp11} is in the fact that it suggests,
in perspective, a general scenario common to all molecules
characterized by $\beta$-sheet-like or $\alpha$-helix-like
conformations and how an electric field is likely to modify it.
}
The probability flux of $\period =40$~ps is quantitatively
comparable to the case $\period =10$~ps, except that the probability fluxes
going into conformation $C$ is stronger, which actually results in
a higher steady probability in $C$.
The probability flux of case $\period =200$~ps is qualitatively
similar to $\period =10$~ps and $\period =40$~ps, however,
the strength is much lower than the latter two cases.
Moreover, 
  we see that the integrated fluxes $Q_{A_2,A_1}$ and $Q_{B_1,A_2}$ reach
  a steadily
increasing stage after about 3200~ps, which is even longer than the time scale
at which the probability of $C$ reaches its steady value. This indicates
a longer intrinsic time scale in the $\period =200$~ps case.
